---
title: "P2"
author: "Robin Singh"
date: "8/24/2021"
output: html_document
---

```{r}
mydata <- read.csv(file="ames_housing_data.csv",head=TRUE,sep=",")
mydata
```

```{r}
mydata$TotalFloorSF <- mydata$FirstFlrSF + mydata$SecondFlrSF
mydata$HouseAge <- mydata$YrSold - mydata$YearBuilt
mydata$QualityIndex <- mydata$OverallQual * mydata$OverallCond
mydata$logSalePrice <- log(mydata$SalePrice)
mydata$price_sqft <- mydata$SalePrice/mydata$TotalFloorSF
summary(mydata$price_sqft)
hist(mydata$price_sqft)
subdata <- subset(mydata, select=c("TotalFloorSF","HouseAge","QualityIndex",
                                  "price_sqft", "SalePrice","LotArea",
                                  "BsmtFinSF1","Neighborhood","HouseStyle",
                                  "LotShape","OverallQual","logSalePrice",
                                  "TotalBsmtSF","HouseStyle","Zoning","LotShape","SaleCondition","Functional", "LotArea","SubClass","LotFrontage","OverallCond", "YearBuilt", "ExterQual", "ExterCond", "FirstFlrSF", "SecondFlrSF", "BedroomAbvGr", "TotRmsAbvGrd", "GrLivArea", "MiscVal", "YearRemodel"))



subdata
```
```{r}
subdatnum <- subset(mydata, select=c("TotalFloorSF","HouseAge","QualityIndex",
                                     "SalePrice","LotArea","OverallQual"))

require(lattice)
pairs(subdatnum, pch = 21)

require(corrplot)
mcor <- cor(subdatnum)
corrplot(mcor, method="shade", shade.col=NA, tl.col="black",tl.cex=0.5)
```
Overall quality and SalePrice seems to have a bit of a linear relationship, as well as quality index but that is to be expected. 

SalePrice and TotalFLoorSF seems to have a slight linear relationship.

logSalePrice and HouseAge seem to have an opposite linear coordination. 

```{r}
library(ggplot2)
ggplot(subdata, aes(x=TotalFloorSF, y=QualityIndex)) + 
  geom_point(color="blue", shape=1) +
  ggtitle("Scatter Plot of Total Floor SF vs QualityIndex") +
  theme(plot.title=element_text(lineheight=0.8, face="bold", hjust=0.5))

ggplot(subdata, aes(x=TotalFloorSF, y=HouseAge)) + 
  geom_point(color="blue", shape=1) +
  ggtitle("Scatter Plot of Total Floor SF vs HouseAge") +
  theme(plot.title=element_text(lineheight=0.8, face="bold", hjust=0.5))

ggplot(subdata, aes(x=TotalFloorSF, y=SalePrice)) + 
  geom_point(color="blue", shape=1) +
  ggtitle("Scatter Plot of Total Floor SF vs QualityIndex") +
  theme(plot.title=element_text(lineheight=0.8, face="bold", hjust=0.5))

ggplot(subdata, aes(x=SalePrice, y=QualityIndex)) + 
  geom_point(color="blue", shape=1) +
  ggtitle("Scatter Plot of Total Floor SF vs HouseAge") +
  theme(plot.title=element_text(lineheight=0.8, face="bold", hjust=0.5))
```
```{r}
subdata = subset(subdata, select = -c(OverallQual,QualityIndex, logSalePrice, price_sqft) )
model1=lm(SalePrice~., data=subdata)
summary(model1)
model1
```
```{r}
model2=lm(SalePrice~TotalFloorSF+HouseAge+LotArea+MiscVal, data=subdata)
summary(model2)
model2
```

```{r}


model3=lm(SalePrice~TotalFloorSF+HouseAge+LotArea+MiscVal, data=subdata)
summary(model3)
model3
```
```{r}
num_cols <- unlist(lapply(subdata, is.numeric))

subdata <- subdata[ ,num_cols]
model4=lm(SalePrice~., data=subdata)
summary(model4)
model4
```
```{r}
subdata=na.omit(subdata)


subdata= subdata[,!(names(subdata) %in% c('LotArea.1', 'SecondFlrSF' ))]
```
```{r}


model5=lm(SalePrice~., data=subdata)
summary(model5)
model5
```

```{r}

mydata$TotalFloorSF <- mydata$FirstFlrSF + mydata$SecondFlrSF
mydata$HouseAge <- mydata$YrSold - mydata$YearBuilt
mydata$QualityIndex <- mydata$OverallQual * mydata$OverallCond
mydata$logSalePrice <- log(mydata$SalePrice)
mydata$price_sqft <- mydata$SalePrice/mydata$TotalFloorSF
subdatawithNeighborhood=mydata
num_cols <- unlist(lapply(subdatawithNeighborhood, is.numeric))

subdatawithNeighborhood <- subdatawithNeighborhood[ ,num_cols]
subdatawithNeighborhood$Neighborhood=mydata$Neighborhood
subdatawithNeighborhood
```


```{r}
subdatawithNeighborhood=na.omit(subdatawithNeighborhood)
subdatawithNeighborhood
subdatawithNeighborhood = subset(subdatawithNeighborhood, select = -c(OverallQual,QualityIndex, logSalePrice, price_sqft, TotalBsmtSF,GrLivArea, MoSold, YrSold,  EnclosedPorch, ThreeSsnPorch) )
model6=lm(SalePrice~., data = subdatawithNeighborhood)
model6
summary(model6)

```









```{r}
model7=lm(SalePrice~TotalFloorSF+HouseAge+LotArea+MiscVal+Neighborhood, data=subdatawithNeighborhood)
model7
summary(model7)
```




```{r}
library('fastDummies')
subdatawithNeighborhood <- dummy_cols(subdatawithNeighborhood, select_columns = 'Neighborhood',
           remove_selected_columns = TRUE)
subdatawithNeighborhood=scale(subdatawithNeighborhood)
corrmatrix <- rcorr(as.matrix(subdatawithNeighborhood))
corrmatrix
```
```{r}
is.matrix(corrmatrix$r)
corrmatrix=corrmatrix$r
isSymmetric(corrmatrix)
```
```{r}
library(corrplot)
corrplot(corrmatrix)
```
```{r}
library(psych)
fa(corrmatrix, nfactors=8) 
#factors_data <- factanal(covmat=corrmatrix, n.obs=2439, factors=3, rotation='varimax');

```
```{r}
fa.parallel(corrmatrix, 
            n.obs=2258, 
            fa="both", 
            n.iter=100,
            fm="ml",
            show.legend=TRUE, main="Scree plot with parallel analysis")
```
```{r}
eigen(corrmatrix)
```
```{r}
factors_data <- fa(r = corrmatrix, nfactors = 15)

#factors_data <- factanal(covmat=corrmatrix, n.obs=2258, factors=15, rotation='varimax', fm='ml');

factors_data
```



```{r}
###########################
# Helper functions
###########################

# find the graph node number by the coordinates
find_node_by_coordinates <- function(x, y, grid_width) {
 return(((y * grid_width) + x) - grid_width)
}

# return the number of observations represented by each node
get_node_counts <- function(x) {
 df <- data.frame(node = x)
 counts <- df %>%
   group_by(node) %>%
   summarize(observations = n())
}

# guideline for grid size = 5 * sqrt(N)
# where N is the number of observations in the data set
find_grid_size <- function(N) {
 return(floor(sqrt(sqrt(N) * 5)))
}

# Shane Lynn 14-01-2014 used to define the palette
coolBlueHotRed <- function(n, alpha = 1) {
 rainbow(n, end=4/6, alpha=alpha)[n:1]
}
```
```{r}
#names(as.data.frame(subdatawithNeighborhood))
subdatawithNeighborhood2=as.data.frame(subdatawithNeighborhood[,0:34])
```

```{r}
subdatawithNeighborhood
```


```{r}

mydata$TotalFloorSF <- mydata$FirstFlrSF + mydata$SecondFlrSF
mydata$HouseAge <- mydata$YrSold - mydata$YearBuilt
mydata$QualityIndex <- mydata$OverallQual * mydata$OverallCond
mydata$logSalePrice <- log(mydata$SalePrice)
mydata$price_sqft <- mydata$SalePrice/mydata$TotalFloorSF
subdatawithNeighborhood2=mydata
num_cols <- unlist(lapply(subdatawithNeighborhood2, is.numeric))

subdatawithNeighborhood2 <- subdatawithNeighborhood2[ ,num_cols]
subdatawithNeighborhood2$Neighborhood=mydata$Neighborhood
subdatawithNeighborhood2
```


```{r}
subdatawithNeighborhood2=na.omit(subdatawithNeighborhood2)
subdatawithNeighborhood2
subdatawithNeighborhood2 = subset(subdatawithNeighborhood2, select = -c(OverallQual,QualityIndex, logSalePrice, price_sqft, TotalBsmtSF,GrLivArea, MoSold, YrSold,  EnclosedPorch, ThreeSsnPorch) )



```
```{r}
require(tidyverse)
require(kohonen)
require(ggplot2)
require(ggridges)


subdatawithNeighborhood2 %>%

 gather(variable, value, -Neighborhood) %>%
 ggplot(aes(y = as.factor(variable),
             fill = Neighborhood,
             x = percent_rank(value))) +
 geom_density_ridges() +
 ggtitle('Class Distributions') +
 theme(plot.title = element_text(hjust = 0.5)) +
 xlab('Value') +
 ylab('Variable')
```
```{r}
scaleddatawneighborhood=as.data.frame(scale(subdatawithNeighborhood2[,-ncol(subdatawithNeighborhood2)]))
scaleddatawneighborhood$Neighborhood=subdatawithNeighborhood2$Neighborhood
scaleddatawneighborhood=scaleddatawneighborhood[,-c(1,2)]
scaleddatawneighborhood
scaleddatawneighborhood %>%

 gather(variable, value, -Neighborhood) %>%
 ggplot(aes(y = as.factor(variable),
             fill = Neighborhood,
             x = percent_rank(value))) +
 geom_density_ridges() +
 ggtitle('Class Distributions') +
 theme(plot.title = element_text(hjust = 0.5)) +
 xlab('Value') +
 ylab('Variable')
```
```{r}
# find all columns having numeric data
numerics <- scaleddatawneighborhood %>%
 select_if(is.numeric) %>%
 names

scaleddatawneighborhood$Neighborhood <- as.factor(scaleddatawneighborhood$Neighborhood)

# find all columns having factors
factors <- scaleddatawneighborhood %>%
 select_if(is.factor) %>%
 names

numerics

```


```{r}
factors
```

```{r}

data_list = list()
distances = vector()

# create a layer for each factor
for (fac in factors){
 data_list[[fac]] = kohonen::classvec2classmat( scaleddatawneighborhood[[fac]] )
 distances = c(distances, 'tanimoto')
}

data_list[['numerics']] = scale(scaleddatawneighborhood[,numerics])
distances = c(distances, 'sumofsquares')
```

```{r}
# review data_list
str(data_list)
names(data_list)

# review distance measures
distances
```
```{r}
# call the find_grid_size function to calculate the grid size
map_dimension = find_grid_size(dim(scaleddatawneighborhood)[1])

# set the number of times the model will cycle through the observations
epochs = 2000
set.seed(123)

# create a grid onto which the som will be mapped
som_grid = somgrid(xdim = 7
                   ,ydim = 7
                   ,topo = "rectangular")


# train the SOM
cc_som = supersom(data_list
                  ,grid = som_grid
                  ,rlen = epochs
                  ,alpha = c(0.1, 0.01)
                  ,whatmap = c(factors,'Neighborhood', 'numerics')
                  ,dist.fcts = distances
                  ,keep.data = TRUE
)

```

dplyr::summarize
library(tidyverse)
plot(cc_som, type = "changes")

plot(cc_som, type = "counts", palette.name = coolBlueHotRed)

cc_som$unit.classif
observations_by_node <- get_node_counts(cc_som$unit.classif)

round(mean(observations_by_node$observations), 2)

plot(cc_som, type = "dist.neighbours", palette.name = coolBlueHotRed)

plot(cc_som, type = "codes", palette.name = coolBlueHotRed)



```{r}
plot(cc_som, type = "changes")
```
```{r}
plot(cc_som, type = "counts", palette.name = coolBlueHotRed)
```
```{r}
cc_som$unit.classif
observations_by_node <- get_node_counts(cc_som$unit.classif)
```
```{r}
plot(cc_som, type = "dist.neighbours", palette.name = coolBlueHotRed)

```
```{r}
anomalies <- find_node_by_coordinates(7, 4, 7)
rows <- which(cc_som$unit.classif == anomalies)
scaleddatawneighborhood[rows, ]
```
```{r}
anomalies <- find_node_by_coordinates(1, 5, 1)
rows <- which(cc_som$unit.classif == anomalies)
scaleddatawneighborhood[rows, ]
```



```{r}
plot(cc_som, type = "codes", palette.name = coolBlueHotRed)
```

```{r}
plot(cc_som, type = "property", property = getCodes(cc_som)[["Neighborhood"]],
main=colnames(getCodes(cc_som))[["Neighborhood"]], palette.name=coolBlueHotRed)
```
```{r}
cc_som$codes[["Neighborhood"]]
```

```{r}
plot(cc_som, type = "quality", palette.name = coolBlueHotRed)
```
```{r}
###########################
# Clustering
###########################

library(RColorBrewer)
n <- 60
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

# fuse all layers into one dataframe
codes = tibble( layers = names(cc_som$codes)
                ,codes = cc_som$codes ) %>%
 mutate( codes = purrr::map(codes, as_tibble) ) %>%
 spread( key = layers, value = codes) %>%
 apply(1, bind_cols) %>%
 .[[1]] %>%
 as_tibble()

# generate distance matrix for codes
dist_m = dist(codes) %>%
 as.matrix()

# generate seperate distance matrix for map location
dist_on_map = kohonen::unit.distances(som_grid)

#exponentiate euclidean distance by distance on map
dist_adj = dist_m ^ dist_on_map

clust_adj = hclust(as.dist(dist_adj), 'ward.D2')

som_cluster_adj = cutree(clust_adj, 11)

plot(cc_som, type="codes", main = "Clusters", bgcol = col_vector[som_cluster_adj], pchs = NA)
```
```{r}
cc_som$unit.classif
observations_by_node <- get_node_counts(cc_som$unit.classif)
```
```{r}

cc.som = supersom(data_list
                  ,grid = som_grid
                  ,rlen = epochs
                  ,alpha = c(0.1, 0.01)
                  ,whatmap = c(factors, 'numerics')
                  ,dist.fcts = distances
                  ,keep.data = TRUE
)

plot(cc.som, type = "changes")

plot(cc.som, type = "counts", palette.name = coolBlueHotRed)

cc.som$unit.classif
observations_by_node <- get_node_counts(cc.som$unit.classif)

round(mean(observations_by_node$observations), 2)

plot(cc.som, type = "dist.neighbours", palette.name = coolBlueHotRed)

plot(cc.som, type = "codes", palette.name = coolBlueHotRed)
```
```{r}
find_grid_size(dim(scaleddatawneighborhood)[1])
```












